---
title: "Marketing Data"
---

```{r setup, include = FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, warning = TRUE, message = FALSE,
                     echo = TRUE, dpi = 180, fig.width = 8, fig.height = 5)
library(tidyverse)
library(silgelib)
library(mice)
library(gridExtra)
library(stargazer)
library(MASS)
library(pscl)
library(lmtest)
library(ztable)
theme_set(theme_bw())
```

# Data Marketing Analysis from Kaggle

## Section 01: Exploratory Data Analysis

Let's analyze the marketing data set provided by an user of Kaggle.

```{r}
setwd("C:/Users/Daniel Gutierrez/Desktop/R Practice/Marketing Data Kaggle")
marketing_data <- readr::read_csv("marketing_data.csv")
```
Simple summary statistics of the data sets to see what variables we have
```{r}
summary(marketing_data)
```
```{r}
str(marketing_data)
```
ID: Customer's unique identifier 


Year_Birth: Customer's birth year

Education: Customer's education level

Marital_Status: Customer's marital status

Income: Customer's yearly household income

Kidhome: Number of children in customer's household

Teenhome: Number of teenagers in customer's household

Dt_Customer: Date of customer's enrollment with the company

Recency: Number of days since customer's last purchase

MntWines: Amount spent on wine in the last 2 years

MntFruits: Amount spent on wine in the last 2 years

MntMeatProducts: Amount spent on meat in the last 2 years

MntFishProducts: Amount spent on fish in the last 2 years

MntSweetProducts: Amount spent on sweets in the last 2 years

MntGoldProds: Amount spent on gold in the last 2 years

NumDealsPurchase: Number of purchases made with a discount

NumWebPurchases: Number of purchases made through the company's web site

NumCatalogPurchases: Number of purchases made using a catalogue

NumStorePurchases: Number of purchases made directly in stores

NumWebVisitsMonth: Number of visits to company's web site in the last month

AcceptedCmp3: 1 if customer accepted the offer in the 3rd campaign, 0 otherwise

AcceptedCmp4: 1 if customer accepted the offer in the 4th campaign, 0 otherwise

AcceptedCmp5: 1 if customer accepted the offer in the 5th campaign, 0 otherwise

AcceptedCmp1: 1 if customer accepted the offer in the 1st campaign, 0 otherwise

AcceptedCmp2: 1 if customer accepted the offer in the 2nd campaign, 0 otherwise

Response: 1 if customer accepted the offer in the last campaign, 0 otherwise

Complain: 1 if customer complained in the last 2 years, 0 otherwise

Country: Customer's location


Note that we have a variable, "year_birth", for the costumers which probably is hard to interpret in a model. We would have to build an age variable. Also, note that the income variable is a string variable. It includes the dollar symbol, "$", before the yearly income. Dt_Customer is a supposed to be a date. 

We have to transform the income variable into a numerical variable. Marital status, education, accepted* and response  as factors. Maybe country as a factor too. Also an age variable: 2021 - year_birth. 

```{r}
marketing <- marketing_data %>% 
  mutate(Income = as.numeric(gsub("[^0-9.-]", "",Income)),
         Dt_Customer = as.Date(Dt_Customer, "%m/%d/%y"),
         Education = as.factor(Education),
         Marital_Status = as.factor(Marital_Status),
         Country = as.factor(Country),
         AcceptedCmp1 = as.factor(AcceptedCmp1),
         AcceptedCmp2 = as.factor(AcceptedCmp2),
         AcceptedCmp3 = as.factor(AcceptedCmp3),
         AcceptedCmp4 = as.factor(AcceptedCmp4),
         AcceptedCmp5 = as.factor(AcceptedCmp5),
         Response = as.factor(Response),
         Age = 2021 - Year_Birth)
marketing
```
The raw the data set now looks better. But there are some NA's. Let's investigate them.
```{r}
summary(marketing)
```
There are 24 NA's for the income variable. We could substitute them by the mean of the income depending on their education. We could also delete them. Let's try both approaches and compare the results.
```{r}
marketing_imp <- marketing %>% 
  group_by(Education) %>% 
  mutate(Income = ifelse(is.na(Income), mean(Income, na.rm = TRUE), Income))
summary(marketing_imp)
```
No more NA's. Now we can start an uni-variate analysis to check for implausible values and outliers. Age and income are already causing me suspicious. Maximum value of Age is 128 years old with a mean of 52 and a 75% quantile of 62, probably this person of 128 is an outlier or just a mistake when writing the dataset. For income the maximum value is \$666,666 per year, with a mean of \$52,254 and 75% of the participants earn less than \$68,290. The maximum value is about 10 times higher. The number itself its already suspicious as it is a sequence of six sixes. It is very likely that this is not a real person. But we analyze it.
```{r}
hist_age <- marketing_imp %>% 
  ggplot(aes(Age))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Age Histogram")+
  ylab(NULL)
hist_age
```


```{r}
sum(marketing_imp$Age > 100)
```
Three participants older than 100 years old. Since this is a digital commerce, I will drop these observations. It is unlikely that the people in this age buy online. 

```{r}
hist_income <- marketing_imp %>% 
  ggplot(aes(Income))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Income Histogram")+
  ylab(NULL)
require(scales)
hist_income + scale_x_continuous(labels = comma)
```
```{r}
sum(marketing_imp$Income > 100000 )
```
There are 13 participants that earn more than $100,000 dollars per year. I will also drop these observations as they might induce noise to the analyses.

```{r}
marketing_ready <- marketing_imp %>% 
  filter(Income < 100000 & Age < 100)
summary(marketing_ready)
```
```{r}
marketing_ready %>% 
  ggplot(aes(Age))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Age Histogram")+
  ylab(NULL)
```


```{r}
marketing_ready %>% 
  ggplot(aes(Income))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Income Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

```

Both variables look better now. Now lets look at the distribution of the remaining continous variables and categorical variables. 
```{r}
hist_wines <- marketing_ready %>% 
  ggplot(aes(MntWines))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Wine Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

hist_fruits <- marketing_ready %>% 
  ggplot(aes(MntFruits))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Fruits Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

hist_MeatProducts <- marketing_ready %>% 
  ggplot(aes(MntMeatProducts))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Meat Products Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

hist_FishProducts <- marketing_ready %>% 
  ggplot(aes(MntFishProducts))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Fish Products Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

hist_SweetProducts <- marketing_ready %>% 
  ggplot(aes(MntSweetProducts))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Sweet Products Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

hist_GoldProds <- marketing_ready %>% 
  ggplot(aes(MntGoldProds))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Gold Products Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

grid.arrange(hist_wines, hist_fruits, hist_MeatProducts, hist_FishProducts,
             hist_SweetProducts, hist_GoldProds, ncol=3)

```
There are a lot of zeros for the distribution of the amount spend on each of the products. A few outliers in the expenditure behavior of the costumers. For example, the meat products histogram shows a point at the right of the distribution with more than 1,500 USD spent in the last two years. The sweet products histogram shows a similar behavior. The gold products are more sold, they have at least higher number of observations with one dollar spent.

```{r}
par(mfrow = c(3,3))
plot(marketing_ready$Education, main = "Education")
plot(marketing_ready$Marital_Status, main = "Marital Status")
plot(marketing_ready$Country, main = "Country")
plot(marketing_ready$AcceptedCmp1, main = "Offer 1st Campaign Success")
plot(marketing_ready$AcceptedCmp2, main = "Offer 2nd Campaign Success")
plot(marketing_ready$AcceptedCmp3, main = "Offer 3rd Campaign Success")
plot(marketing_ready$AcceptedCmp4, main = "Offer 4th Campaign Success")
plot(marketing_ready$AcceptedCmp1, main = "Offer 5th Campaign Success")
plot(marketing_ready$Response, main = "Last offer Success")
```
```{r}
table(marketing_ready$Marital_Status)
```
Probably would make sense to merge Alone with the Single category and drop "Absurd" and "YOLO". 

```{r}
marketing_ready$Marital_Status <- as.factor(ifelse(as.character(marketing_ready$Marital_Status) == "Alone", "Single",as.character(marketing_ready$Marital_Status)))
table(marketing_ready$Marital_Status)
```
```{r}
marketing_ready <- marketing_ready %>% 
  filter(Marital_Status != "Absurd" & Marital_Status != "YOLO") %>% 
  mutate(Marital_Status = factor(Marital_Status))
summary(marketing_ready)
```
We could also create a variable for total purchases in the last two years
```{r}
marketing_ready <- marketing_ready %>% 
  mutate(total_purch = MntWines + MntFruits + MntMeatProducts + MntFishProducts + MntSweetProducts + MntGoldProds)
marketing_ready %>% 
  ggplot(aes(total_purch)) +
  geom_histogram()
```
```{r}
summary(marketing_ready)
```
```{r}
marketing_ready %>% 
  ggplot(aes(Income, log(total_purch)))+
  geom_point()+
  geom_smooth()
```
There is a non-linear relationship between amount spent and income.

### Section 02: Statistical Analysis

1. What factors are significantly related to the number of store purchases?
2. Does US fare significantly better than the Rest of the World in terms of total purchases?
3. Your supervisor insists that people who buy gold are more conservative. Therefore, people who spent an above average amount on gold in the last 2 years would have more in store purchases. Justify or refute this statement using an appropriate statistical test
4. Fish has Omega 3 fatty acids which are good for the brain. Accordingly, do "Married PhD candidates" have a significant relation with amount spent on fish? What other factors are significantly related to amount spent on fish? (Hint: use your knowledge of interaction variables/effects)
5. Is there a significant relationship between geographical regional and success of a campaign?

To answer question 1 we can regress the number of store purchases on age, education, marital status, income, number of kids and teenagers at home and country. Also a second regression could include the offers the costumer took. A third regression could also include an indicator variable for each type of product to identify which products are acquired in a store. It is interesting to investigate a count model that estimates the probabilities of the costumers purchasing in a store.

```{r}
marketing_ready <- marketing_ready %>% 
  mutate(wine = factor(ifelse(MntWines > 0, 1, 0)),
         fruits = factor(ifelse(MntFruits > 0, 1, 0)),
         meat = factor(ifelse(MntMeatProducts > 0, 1, 0)),
         fish = factor(ifelse(MntFishProducts > 0, 1, 0)),
         sweet = factor(ifelse(MntSweetProducts > 0, 1, 0)),
         gold = factor(ifelse(MntGoldProds > 0, 1, 0)),
         store_purchase = factor(ifelse(NumStorePurchases > 0, 1,0)))

store_ols1 <- lm(NumStorePurchases ~ Age + Education + Marital_Status + Income + Kidhome + Teenhome + Country, data = marketing_ready)
store_ols2 <- lm(NumStorePurchases ~ Age + Education + Marital_Status + Income + Kidhome + Teenhome + Country + AcceptedCmp1 + AcceptedCmp2 + AcceptedCmp3 + AcceptedCmp4 + AcceptedCmp5 + Response, data = marketing_ready)
store_ols3 <- lm(NumStorePurchases ~ Age + Education + Marital_Status + Income + Kidhome + Teenhome + Country + AcceptedCmp1 + AcceptedCmp2 + AcceptedCmp3 + AcceptedCmp4 + AcceptedCmp5 + Response + wine + fruits + meat + fish + sweet + gold, data = marketing_ready)
store_negbin1 <- glm(NumStorePurchases ~ Age + Education + Marital_Status + Income + Kidhome + Teenhome + Country + AcceptedCmp1 + AcceptedCmp2 + AcceptedCmp3 + AcceptedCmp4 + AcceptedCmp5 + Response + wine + fruits + meat + fish + sweet + gold, family = "poisson", data = marketing_ready)
stargazer(store_ols1, store_ols2, store_ols3, store_negbin1, type = "text", align = TRUE, no.space = TRUE)
```
The OLS models hint us that the factors driven store purchases are kids and teenagers at home, income, the offer accepted at the moment of the purchase and the product that the costumer buys. The count model confirms these findings. It seems that the second and fourth offer affects positively purchasing in stores. Fish and sweets are the products that also contributes to more purchases. Income follows the same pattern, and surprisingly age contributes negatively to store purchases. 


To answer question 2, lets plot a graph to observe how the country contribution to total purchases.
```{r}
marketing_ready %>% 
  ggplot(aes(Country, total_purch, fill = Country)) +
  geom_col() +
  scale_y_continuous() +
  coord_flip() +
  scale_fill_brewer(palette = "Spectral")
```
By comparing only countries total purchases, Spain is doing better than any other country. However, Spain is also the country with more costumers.
```{r}
summary(marketing_ready$Country)
```

Lets use a boxplot to investigate this,
```{r}
marketing_ready %>% 
  ggplot(aes(total_purch, Country, fill = Country)) +
  geom_boxplot(outlier.alpha = 0.5) +
  scale_x_continuous() +
  scale_fill_brewer(palette = "Spectral")
```
Now we can see that the median of the US is higher than SP. ME is by far the country with more total purchases. 
```{r}
marketing_ready %>% 
  group_by(Country) %>% 
  mutate(mean_country = mean(total_purch)) %>% 
  ggplot(aes(Country, mean_country, colour = Country)) +
  geom_point() +
  scale_colour_brewer(palette = "Spectral")

```
In average, the US has a higher total purchases value than SP, IND and AUS but GER and CA are slightly better than US, and ME is better. We can run a regression to see these effects.

```{r}
marketing_country <- marketing_ready %>% 
  mutate(Country = relevel(Country, ref = "US"))

totalpurch1 <- lm(log(total_purch) ~ Age + Income + I(Income^2) + Country + Kidhome + Teenhome + AcceptedCmp1 + AcceptedCmp1 + AcceptedCmp2 + AcceptedCmp3 + AcceptedCmp4 + AcceptedCmp5 + Response + Complain, data = marketing_country)
stargazer(totalpurch1, type = "text", aling = TRUE, no.space = TRUE)
```
What we have found with the regression analysis is that average total purchases in AUS, GER and SP are lower than in the US. In average, AUS total purchases are 21.8% lower compared to the US. Same pattern for GER and SP, 27.8% and 16.9% lower.

To answer question, we create an indicator variable that equals 1 whether the amount spent on gold is higher than the average or not. Then regress the number of store purchases on this indicator variable. The mean of the amount spend on gold is 43.81. 

```{r}
summary(marketing_ready$MntGoldProds)
```

```{r}
marketing_ready <- marketing_ready %>% 
  mutate(gold_above = as.factor(ifelse(MntGoldProds > 45, 1, 0)))
summary(marketing_ready$gold_above)

above1 <- lm(NumStorePurchases ~ Age + Education + Marital_Status + Income + Kidhome + Teenhome + Country + AcceptedCmp1 + AcceptedCmp2 + AcceptedCmp3 + AcceptedCmp4 + AcceptedCmp5 + Response + gold_above, data = marketing_ready)
stargazer(above1, type = "text", no.space = TRUE, align = TRUE)
```
The result of the regression confirms the supervisor's hypothesis. Among gold buyers, those who buy more than the average go more often to the store.

Lets now analyze question 4. To investigate that, I will run a regression interacting marital status and education. To answer it correctly, the reference category for marital status is "married" and for education "PhD". In this sense, we can compare the categories with the other ones.

```{r}
marketing_ready <- marketing_ready %>% 
  mutate(Marital_Status = relevel(Marital_Status, ref = "Married"),
         Education = relevel(Education, ref = "PhD"))
inter1 <- lm(MntFishProducts ~ 0 + Age + Education + Marital_Status + Education*Marital_Status + Income + I(Income^2) + Kidhome + Teenhome + Country + AcceptedCmp1 + AcceptedCmp2 + AcceptedCmp3 + AcceptedCmp4 + AcceptedCmp5 + Response, data = marketing_ready)
stargazer(inter1, type = "text", no.space = TRUE, align = TRUE)
```
The results of the regression tell us that married PhD candidates only buy more fish than graduated divorced participants. Widow participants tend to spend more in fish products. In general, divorced participants buy more fish too. Kids and teens at home decrease the amount spent in fish. Those who accepted offer campaign 2 buy more fish. Accepting campaigns 4 and5 spend less in fish than those who did not accept these campaigns. 

To answer question 5, we can explore graphically the different campaigns.
```{r}
par(mfrow = c(2,3))
plot(marketing_ready$AcceptedCmp1 ~ marketing_ready$Country, main = "Campaing 1", ylab = "Accepted Offer", xlab = "Country")
plot(marketing_ready$AcceptedCmp2 ~ marketing_ready$Country, main = "Campaing 2", ylab = NULL, xlab = "Country")
title(ylab = NULL)
plot(marketing_ready$AcceptedCmp3 ~ marketing_ready$Country, main = "Campaing 3", ylab = NULL, xlab = "Country")
plot(marketing_ready$AcceptedCmp4 ~ marketing_ready$Country, main = "Campaing 4", ylab = "Accepted Offer", xlab = "Country")
plot(marketing_ready$AcceptedCmp5 ~ marketing_ready$Country, main = "Campaing 5", ylab = NULL, xlab = "Country")
plot(marketing_ready$Response ~ marketing_ready$Country, main = "Last Campaign", ylab = NULL, xlab = "Country")
```

