---
title: "Marketing Data"
output: html_notebook
---

```{r setup, include = FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, warning = TRUE, message = FALSE,
                     echo = TRUE, dpi = 180, fig.width = 8, fig.height = 5)
library(tidyverse)
library(silgelib)
library(mice)
library(gridExtra)
theme_set(theme_bw())
```

# Data Marketing Analysis from Kaggle

## Section 01: Exploratory Data Analysis

Let's analyze the marketing data set provided by an user of Kaggle.

```{r}
setwd("C:/Users/Daniel Gutierrez/Desktop/R Practice/Marketing Data Kaggle")
marketing_data <- readr::read_csv("marketing_data.csv")
```
Simple summary statistics of the data sets to see what variables we have
```{r}
summary(marketing_data)
```
```{r}
str(marketing_data)
```
ID: Customer's unique identifier
Year_Birth: Customer's birth year
Education: Customer's education level
Marital_Status: Customer's marital status
Income: Customer's yearly household income
Kidhome: Number of children in customer's household
Teenhome: Number of teenagers in customer's household
Dt_Customer: Date of customer's enrollment with the company
Recency: Number of days since customer's last purchase
MntWines: Amount spent on wine in the last 2 years
MntFruits: Amount spent on wine in the last 2 years
MntMeatProducts: Amount spent on meat in the last 2 years
MntFishProducts: Amount spent on fish in the last 2 years
MntSweetProducts: Amount spent on sweets in the last 2 years
MntGoldProds: Amount spent on gold in the last 2 years
NumDealsPurchase: Number of purchases made with a discount
NumWebPurchases: Number of purchases made through the company's web site
NumCatalogPurchases: Number of purchases made using a catalogue
NumStorePurchases: Number of purchases made directly in stores
NumWebVisitsMonth: Number of visits to company's web site in the last month
AcceptedCmp3: 1 if customer accepted the offer in the 3rd campaign, 0 otherwise
AcceptedCmp4: 1 if customer accepted the offer in the 4th campaign, 0 otherwise
AcceptedCmp5: 1 if customer accepted the offer in the 5th campaign, 0 otherwise
AcceptedCmp1: 1 if customer accepted the offer in the 1st campaign, 0 otherwise
AcceptedCmp2: 1 if customer accepted the offer in the 2nd campaign, 0 otherwise
Response: 1 if customer accepted the offer in the last campaign, 0 otherwise
Complain: 1 if customer complained in the last 2 years, 0 otherwise
Country: Customer's location


Note that we have a variable, "year_birth", for the costumers which probably is hard to interpret in a model. We would have to build an age variable. Also, note that the income variable is a string variable. It includes the dollar symbol, "$", before the yearly income. Dt_Customer is a supposed to be a date. 

We have to transform the income variable into a numerical variable. Marital status, education, accepted* and response  as factors. Maybe country as a factor too. Also an age variable: 2021 - year_birth. 

```{r}
marketing <- marketing_data %>% 
  mutate(Income = as.numeric(gsub("[^0-9.-]", "",Income)),
         Dt_Customer = as.Date(Dt_Customer, "%m/%d/%y"),
         Education = as.factor(Education),
         Marital_Status = as.factor(Marital_Status),
         Country = as.factor(Country),
         AcceptedCmp1 = as.factor(AcceptedCmp1),
         AcceptedCmp2 = as.factor(AcceptedCmp2),
         AcceptedCmp3 = as.factor(AcceptedCmp3),
         AcceptedCmp4 = as.factor(AcceptedCmp4),
         AcceptedCmp5 = as.factor(AcceptedCmp5),
         Response = as.factor(Response),
         Age = 2021 - Year_Birth)
marketing
```
The raw the data set now looks better. But there are some NA's. Let's investigate them.
```{r}
summary(marketing)
```
There are 24 NA's for the income variable. We could substitute them by the mean of the income depending on their education. We could also delete them. Let's try both approaches and compare the results.
```{r}
marketing_imp <- marketing %>% 
  group_by(Education) %>% 
  mutate(Income = ifelse(is.na(Income), mean(Income, na.rm = TRUE), Income))
summary(marketing_imp)
```
No more NA's. Now we can start an uni-variate analysis to check for implausible values and outliers. Age and income are already causing me suspicious. Maximum value of Age is 128 years old with a mean of 52 and a 75% quantile of 62, probably this person of 128 is an outlier or just a mistake when writing the dataset. For income the maximum value is \$666,666 per year, with a mean of \$52,254 and 75% of the participants earn less than \$68,290. The maximum value is about 10 times higher. The number itself its already suspicious as it is a sequence of six sixes. It is very likely that this is not a real person. But we analyze it.
```{r}
hist_age <- marketing_imp %>% 
  ggplot(aes(Age))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Age Histogram")+
  ylab(NULL)
hist_age
```


```{r}
sum(marketing_imp$Age > 100)
```
Three participants older than 100 years old. Since this is a digital commerce, I will drop these observations. It is unlikely that the people in this age buy online. 

```{r}
hist_income <- marketing_imp %>% 
  ggplot(aes(Income))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Income Histogram")+
  ylab(NULL)
require(scales)
hist_income + scale_x_continuous(labels = comma)
```
```{r}
sum(marketing_imp$Income > 100000 )
```
There are 13 participants that earn more than $100,000 dollars per year. I will also drop these observations as they might induce noise to the analyses.

```{r}
marketing_ready <- marketing_imp %>% 
  filter(Income < 100000 & Age < 100)
summary(marketing_ready)
```
```{r}
marketing_ready %>% 
  ggplot(aes(Age))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Age Histogram")+
  ylab(NULL)
```


```{r}
marketing_ready %>% 
  ggplot(aes(Income))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Income Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

```

Both variables look better now. Now lets look at the distribution of the remaining continous variables and categorical variables. 
```{r}
hist_wines <- marketing_ready %>% 
  ggplot(aes(MntWines))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Wine Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

hist_fruits <- marketing_ready %>% 
  ggplot(aes(MntFruits))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Fruits Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

hist_MeatProducts <- marketing_ready %>% 
  ggplot(aes(MntMeatProducts))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Meat Products Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

hist_FishProducts <- marketing_ready %>% 
  ggplot(aes(MntFishProducts))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Fish Products Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

hist_SweetProducts <- marketing_ready %>% 
  ggplot(aes(MntSweetProducts))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Sweet Products Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

hist_GoldProds <- marketing_ready %>% 
  ggplot(aes(MntGoldProds))+
  geom_histogram(bins = 50, fill = "#24B0B7", alpha = 0.8)+
  labs(title = "Gold Products Histogram")+
  ylab(NULL) + scale_x_continuous(labels = comma)

grid.arrange(hist_wines, hist_fruits, hist_MeatProducts, hist_FishProducts,
             hist_SweetProducts, hist_GoldProds, ncol=3)

```
There are a lot of zeros for the distribution of the amount spend on each of the products. A few outliers in the expenditure behavior of the costumers. For example, the meat products histogram shows a point at the right of the distribution with more than 1,500 USD spent in the last two years. The sweet products histogram shows a similar behavior. The gold products are more sold, they have at least higher number of observations with one dollar spent.

```{r}
par(mfrow = c(3,3))
plot(marketing_ready$Education, main = "Education")
plot(marketing_ready$Marital_Status, main = "Marital Status")
plot(marketing_ready$Country, main = "Country")
plot(marketing_ready$AcceptedCmp1, main = "Offer 1st Campaign Success")
plot(marketing_ready$AcceptedCmp2, main = "Offer 2nd Campaign Success")
plot(marketing_ready$AcceptedCmp3, main = "Offer 3rd Campaign Success")
plot(marketing_ready$AcceptedCmp4, main = "Offer 4th Campaign Success")
plot(marketing_ready$AcceptedCmp1, main = "Offer 5th Campaign Success")
plot(marketing_ready$Response, main = "Last offer Success")
```
```{r}
table(marketing_ready$Marital_Status)
```
Probably would make sense to merge Alone with the Single category and drop "Absurd" and "YOLO". 

```{r}
marketing_ready$Marital_Status <- as.factor(ifelse(as.character(marketing_ready$Marital_Status) == "Alone", "Single",as.character(marketing_ready$Marital_Status)))
table(marketing_ready$Marital_Status)
```
```{r}
marketing_ready <- marketing_ready %>% 
  filter(Marital_Status != "Absurd" & Marital_Status != "YOLO") %>% 
  mutate(Marital_Status = factor(Marital_Status))
summary(marketing_ready)
```
### Section 02: Statistical Analysis

