---
title: "Marketing Data"
output: html_notebook
---

```{r setup, include = FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, warning = TRUE, message = FALSE,
                     echo = TRUE, dpi = 180, fig.width = 8, fig.height = 5)
library(tidyverse)
library(silgelib)
theme_set(theme_plex())
```

# Data Marketing Analysis from Kaggle

Let's analyze the marketing data set provided by an user of Kaggle.

```{r}
setwd("C:/Users/Daniel Gutierrez/Desktop/R Practice/Marketing Data Kaggle")
marketing_data <- readr::read_csv("marketing_data.csv")
```
Simple summary statistics of the data sets to see what variables we have
```{r}
summary(marketing_data)
str(marketing_data)
```
Note that we have a variable, "year_birth", for the costumers which probably is hard to interpret in a model. We would have to build an age variable. Also, note that the income variable is a string variable. It includes the dollar symbol, "$", before the yearly income. Dt_Customer is a supposed to be a date. 

We have to transform the income variable into a numerical variable. Marital status, education, accepted* and response  as factors. Maybe country as a factor too. Also an age variable: year of birth - 2021. 

```{r}

```

